---
title: "Use dictionary"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
#output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualising the gdpm package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This package contains a dictionary for Vietnamese provinces names, districts names and communes names, useful to ensure consistency between different data packages using Vietnamese provinces, districts and/or communes. It also contains dictionaries for the Camdodian, Lao and Thai provinces names.
The package also contains the list names of the provinces of Vietnam ordered by year. One specificity of Vietnam is the splits of administrative provinces during its history, starting from 40 provinces in 1980 and ending with 63 provinces since 2008. This list contains for each of change, the name of the "new" provinces in English.

```{r eval=F}
devtools::install_github("choisy/dictionary")
```

Loading and attaching the `dictionary` package:

```{r}
library(dictionary)
```

# Usage examples

## Dictionnary

We can use some examples to demonstrate our package: 

### Communes

We create a vector containing some communes names in Vietnamese:

```{r}
communes_vn <- c("Xã Nghĩa Lợi", "Xã Nam Mẫu", "Xã Xuân Vinh", "Thị trấn Thứ Ba", "Xã Hải Dương", "Xã Vĩnh Lộc A", "Xã Nội Hoàng", "Xã Điện Trung", "Xã Hoàng Đồng", "Xã Nàn Sán")
```

We can use the `character vector` `communes` to translate the Vietnamese vector but first the vector should be encode in UNICODE to avoid encoding trouble. For that, we use the `stringi` package:

```{r}
uni_communes_vn <- stringi::stri_escape_unicode(communes_vn)
```

And now, we can use the vector `communes`

```{r}
commune_transl <- communes[uni_communes_vn]
```

We obtain a `character vector`, with the original names as `name` and the translation as `value`:   

```{r}
str(commune_transl)
commune_transl
```

### Districts and Provinces

It's the same principle for the provinces and the districts:
We create two vectors that needed to be translate: 

```{r}
districts_vn <- c("Huyện Tây Hòa", "Huyện Kim Thành", "Quận Liên Chiểu", "Huyện Tương Dương", "Huyện Bình Đại", "Thị xã Mường Lay", "Thành phố Vĩnh Long", "Huyện Gia Viễn", "Huyện Văn Quan", "Thị xã Sông Công")
provinces_vn <- c("Ninh Bình", "Hà Nội", "An Giang", "Bình Phước", "Lào Cai", "Phú Thọ", "Bắc Giang", "Vĩnh Phúc", "Bạc Liêu", "Hà Giang")
```

But, we use the `character vector` `districts` and `provinces` contains in the package to translate the two vectors after changing the encoding to UNICODE:

```{r}
uni_districts_vn <- stringi::stri_escape_unicode(districts_vn)
uni_provinces_vn <- stringi::stri_escape_unicode(provinces_vn)
```

```{r}
district_transl <- districts[uni_districts_vn]
province_transl <- vn_province[uni_provinces_vn]
```

We obtain two `character vectors`, with the original names as `name` and the translation as `value`:

```{r}
str(district_transl)
district_transl
str(province_transl)
province_transl
```

## List of province per year

It's a `list` object : 

```{r}
str(province_year)
```

To see a particular year (for example 1997) :

```{r}
province_year$`1997`
```

